### 2017-12-07 tex/petri_net.tex

\centering
\begin{tikzpicture}[x=2cm,y=2cm,>=stealth',bend angle=45,auto]

  \tikzstyle{place}=[circle,thick,draw=blue!75,fill=blue!20,minimum size=6mm]
  \tikzstyle{transition}=[rectangle,thick,draw=black!75,
  			  fill=black!20,minimum size=4mm]

%  \tikzstyle{every label}=[blue!75]

  \begin{scope}
    \node [place,tokens=1] (s1) at (0, 1) [label={[align=center]left:Receive\\ feedback}] {};
    \node [place] (s2) at (1, 2) [label={[align=center]above:Received\\ ACK}] {};
    \node [place] (s3) at (2, 1) [label={[align=center]right:Packets\\ in the\\ network}] {};
    \node [place] (s4) at (1, 0) [label={[align=center]below:Open congestion\\ window (cwnd)}] {};

    \node [transition] (e1) at (0, 0) [label=left:Take action] {}
      edge [post] node {$a$}(s4)
      edge [pre] node {1} (s1);

    \node [transition] (e2) at (2, 0) [label={[align=center]right:Send\\ packet}] {}
      edge [pre] node {1} (s4)
      edge [post] node {1} (s3);

    \node [transition] (e3) at (2, 2) [label={[align=center]right:Receive\\ ACK}] {}
      edge [pre] node {1} (s3)
      edge [post] node {1} (s2);

    \node [transition] (e4) at (1, 1) [label={[align=center]left:Update\\open\\cwnd}] {}
      edge [pre] node {1} (s2)
      edge [post] node {1} (s4);

    \node [transition] (e5) at (0, 2) [label={[align=center]left:Received\\ feedback}] {}
      edge [pre] node {1} (s2)
      edge [post] node {1} (s1);

  \end{scope}

\end{tikzpicture}