### 2018-01-14 tex/neural_network.tex

\centering
\begin{tikzpicture}[every label/.append style={font=\tiny}, node distance=1.2cm, >=stealth',bend angle=45,auto]

  \tikzstyle{place}=[circle,thick,draw=blue!75,fill=blue!20,minimum size=6mm]
  \tikzstyle{input place}=[circle,thick,draw=olive!75,fill=olive!20,minimum size=6mm]
  \tikzstyle{output place}=[circle,thick,draw=magenta!75,fill=magenta!20,minimum size=6mm]
   \tikzstyle{lstm place}=[circle,thick,draw=red!75,fill=red!20,minimum size=6mm]
   \tikzstyle{softplus place}=[circle,thick,draw=green!75,fill=green!20,minimum size=6mm]

  \tikzstyle{transition}=[rectangle,thick,draw=black!75,
  			  fill=black!20,minimum size=4mm]

%  \tikzstyle{every label}=[blue!75]

  \begin{scope}
    \node [input place] (s1) [label={[align=center]left: State (12)}] {};
    \node [place] (s2) [above right of=s1, label={[align=center]above: Linear}] {};

  \path[->] (s1) edge node {} (s2);

    \node [lstm place] (s3) [right of=s2, label={[align=center]below: GRU (32)}] {};

  \path[->] (s2) edge node {} (s3); 
  \path[->] (s3) edge [loop above] node {} (s3);

    \node [lstm place] (s4) [right of=s3, label={[align=center]below: GRU (32)}] {};

  \path[->] (s3) edge node {} (s4);
  \path[->] (s4) edge [loop above] node {} (s4);

    \node [place] (s7) [right of=s4, label={[align=center]below:{ Linear}}] {};

  \path[->] (s4) edge node {} (s7);

    \node [output place] (s8) [right of=s7, label={[align=center]below: $\sigma$~(1)}] {};

  \path[->] (s7) edge node {} (s8);

    \node [place] (s5) [above of=s7, label={[align=center]above: Linear}] {};

  \path[->] (s4) edge node {} (s5);

    \node [output place] (s6) [right of=s5, label={[align=center]above: $\mu$ (1)}] {};

  \path[->] (s5) edge node {} (s6);

    \node [place] (s10) [below right of=s1, label={[align=center]below: Linear}] {};

  \path[->] (s1) edge node {} (s10);

    \node [lstm place] (s11) [right of=s10, label={[align=center]above: GRU (32)}] {};

  \path[->] (s10) edge node {} (s11);
  \path[->] (s11) edge [loop below] node {} (s11);

    \node [lstm place] (s12) [right of=s11, label={[align=center]above: GRU (32)}] {};

  \path[->] (s11) edge node {} (s12);
  \path[->] (s12) edge [loop below] node {} (s12);

    \node [place] (s13) [right of=s12, label={[align=center]above: Linear}] {};

  \path[->] (s12) edge node {} (s13);

    \node [output place] (s17) [right of=s13, label={[align=center]above: $V_\text{packets}$ (1)}] {};

  \path[->] (s13) edge node {} (s17);

    \node [place] (s15) [below of=s13, label={[align=center]above: Linear}] {};

  \path[->] (s12) edge node {} (s15);

    \node [output place] (s19) [right of=s15, label={[align=center]above: $V_\text{delay}$ (1)}] {};

  \path[->] (s15) edge node {} (s19);

    \node [place] (s16) [below of=s15, label={[align=center]above: Linear}] {};

  \path[->] (s12) edge node {} (s16);

    \node [output place] (s20) [right of=s16, label={[align=center]above: $V_\text{sent}$ (1)}] {};

  \path[->] (s16) edge node {} (s20);

    \node [place] (s14) [below of=s16, label={[align=center]above: Linear}] {};

  \path[->] (s12) edge node {} (s14);

    \node [output place] (s21) [right of=s14, label={[align=center]above: $V_\text{duration}$ (1)}] {};

  \path[->] (s14) edge node {} (s21);

%    \node [place] (s17) [below of=s14, label={[align=center]above: Linear}] {};
%    
%  \path[->] (s12) edge node {} (s17);
%    
%    \node [output place] (s22) [right of=s17, label={[align=center]above: $V_\text{byte}$ (1)}] {};   
%    
%  \path[->] (s17) edge node {} (s22);

  \end{scope}

\end{tikzpicture}